### 2018-03-01 inst/doc/MSEtool.Rmd
gravity movement TMB almost working!

---
title: "MSEtool"
subtitle: "Development, Testing and Implementation of Management Procedures for Data-Rich Fisheries (v1.1)"
author: "Tom Carruthers (<t.carruthers@oceans.ubc.ca>) 
         Adrian Hordyk  (<a.hordyk@oceans.ubc.ca>),
         Quang Huynh (<q.huynh@oceans.ubc.ca>)"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
  rmarkdown::pdf_vignette:
    toc: true
    number_sections: true    
vignette: >
  %\VignetteIndexEntry{MSEtool}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<style type="text/css">

body{ /* Normal  */
   font-size: 12px;
}
td {  /* Table  */
   font-size: 8px;
}
h1 { /* Header 1 */
 font-size: 18px;
 color: DarkBlue;
}
h2 { /* Header 2 */
 font-size: 15px;
 color: DarkBlue;
}
h3 { /* Header 3 */
 font-size: 14px;
 color: DarkBlue;
}
code.r{ /* Code block */
  font-size: 10px;
}
pre { /* Code block */
  font-size: 10px
}
</style>

```{r set options, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
knitr::opts_chunk$set(dpi=85)
options(width = 650)
```

<br>

# Foreword

MSEtool development is funded by the Canadian Department of Fisheries and Oceans and benefits from ongoing collaborations with the Natural Resources Defense Council and number of Canadian government scientists including Robyn Forrest, Sean Anderson and Daniel Duplisea.

The package uses the Management Strategy Evaluation engine of [DLMtool](http://www.datalimitedtoolkit.org) to undertake closed-loop simulation testing of management procedures. The DLMtool package was developed at the University of British Columbia in collaboration with the Natural Resources Defense Council.

DLMtool development has been funded by the Gordon and Betty Moore Foundation, the Packard Foundation, U.S. National Oceanic and Atmospheric Administration, Fisheries and Oceans Canada, the Walton Foundation, Resources Legacy Fund, the Natural Resources Defense Council, the United Nations Food and Agricultural Organization and the California Department of Fish and Wildlife.

<br>

# Version Notes

The package is subject to ongoing testing. If you find a bug or a problem please send a report to <t.carruthers@oceans.ubc.ca> so that it can be fixed!

## New Additions to this Version (v1.1)

* The MCMC outputs of iSCAM models are now used to characterise operating model uncertainty

* iSCAM2DLM function added to rip from iSCAM assessments

* SS2DLM function updated for latest version of R4SS

## Coming soon

Data-rich MPs

* Statistical Catch at Age (SCA)
* Age-Structure Production Models (ASPM)
* State-Space Delay Difference models (DD_SS)
* Surplus Production models (SP)
* State-Space Surplus Production Models (SP_SS)
* Kalman filer assessment approximation

Visualization

* Standardized assessment reports

# Introduction

## Management Strategy Evaluation and DLMtool

<img src="DLMtool_logo.png" width="90">

Some text

## The operating model

## MSE design

## The MSE process

<img src="DLMtool_process.jpg" width="400">

<br>

#References

```{r, echo=FALSE, eval=FALSE}
sfStop()                            
```

<br>